\textbf{Analyzing Machupo virus-receptor binding by molecular dynamics simulations}

### 2014-01-20 full_paper.tex
Fixed figure order.

The hTfR1 is the primary receptor used by MACV for binding its host cell prior to infection. The primary role of hTfR1 $in~vivo$ is to bind transferrin for cellular iron uptake. The hTfR1 protein contains three extracellular domains: two basilar domains and an apical domain. The two basilar domains serve most of the transferrin-binding function \citep{Abraham2010,Rad20112}. Viral entry is initiated by GP1 binding to the apical domain of hTfR1. Previous work has indicated that the GP1/hTfR1 binding interaction is the primary determinant of MACV host range variation \citep{Rad20111,Rad20112}. The co-crystal structure shows that the high affinity interaction between GP1 and hTfR1 forces the normally flexible loop in the apical domain of hTfR1 into a rigid $\beta$-pleated sheet domain. For GP1, several extended loops mediate binding to hTfR1 \citep{Abraham2010,Rad20112}, and many of the interface interactions are mediated by extensive hydrogen-bonding networks \citep{Abraham2010} (Figure~\ref{fig:networks}). Experimental alanine-scanning and whole-cell infectivity assays have identified several sites in both GP1 and hTfR1 that are probably critical for establishing infection \citep{Rad20111,Rad20112}.

@@

\subsection{The GP1/hTfR1 system}
The GP1/hTfR1 interface marks a particularly important and useful test system. There are several sites on both the human and viral protein known to affect the infectivity phenotype of MACV. Many of the important sites have been mapped by \textit{in vitro} flow-cytometry based entry assays. The GP1/hTfR1 interface appears not to be dominated by one particular type of interaction (electrostatics, hydrogen-bonding, or van der Waals). In addition, much of the binding domain on hTfR1 is on a loop that is flexible prior to viral binding, but organizes to become a strand of a $\beta$-sheet on binding. As a result,  many other computational techniques \citep{Grant2011,Kortemme2004} are only marginally useful. The complex nature of this interface represents a particularly difficult challenge for traditional computational analysis.

In total, we tested 7 point mutants and 3 double mutants in addition to the WT complex (Table~\ref{tab:summary_mutants} and Figure~\ref{fig:mutant_comparison}). All of the mutations are within 5~\AA\ of the protein--protein interface. Mutations in hTfR1 at site 211 have proven capable of causing loss-of-entry according to \textit{in vitro} flow-cytometry infection assays or known host-range limitations \citep{Rad2008,Rad20111,Rad20112}. Most likely, this effect is caused by the destruction of a critical hydrogen bond to Ser113 or Ser111 in GP1. The lost hydrogen bond would lead to the subsequent loss of a large hydrogen-bonding network seen in the crystal structure (Table~\ref{tab:summary_mutants}) \citep{Abraham2010}. In a manner similar to site 211, N348 appears to be important for binding by participating in a critical hydrogen bonding network \citep{Rad2008,Abraham2010} to GP1. In particular, N348Lys is reported in the literature to cause significantly reduced viral entry \textit{in vivo} (Table~\ref{tab:summary_mutants}) \citep{Rad2008,Abraham2010}. Finally, an alanine mutation at site 111 in GP1 (mutation vR111A) has also been shown to cause decreased entry (Table~\ref{tab:summary_mutants}) \citep{Rad20112}. For notation purposes, the viral site is always referred to with a preceding `v'.

@@

\subsection{Comparative analysis of the GP1/hTfR1 interface}
Considering the involvement of extended hydrogen-bonding networks in the GP1/hTfR1 interface, it was not clear that individual alanine mutations, even those that should destroy such networks, would significantly change the strength of interaction. One major advantage of first principles simulations is the ability to test mutations other than alanine without additional underlying assumptions in the energy function. As shown in Table~\ref{tab:summary_mutants}, we made additional mutations based on biochemical intuition or available experimental data to chemically diverse amino acids including tryptophan, lysine, aspartate, and threonine. Several mutations caused significant relative affinity changes. In addition, to detect synergistic effects, we tested several double mutants where both mutations appeared to cause similar changes in binding. Then, we compared the size of those differences to single mutants (Figure~\ref{fig:sensitivity} and~\ref{fig:mutant_comparison}).

>>>>

The hTfR1 is the primary receptor used by MACV for binding its host cell prior to infection. The primary role of hTfR1 $in~vivo$ is to bind transferrin for cellular iron uptake. The hTfR1 protein contains three extracellular domains: two basilar domains and an apical domain. The two basilar domains serve most of the transferrin-binding function \citep{Abraham2010,Rad20112}. Viral entry is initiated by GP1 binding to the apical domain of hTfR1. Previous work has indicated that the GP1/hTfR1 binding interaction is the primary determinant of MACV host range variation \citep{Rad20111,Rad20112}. The co-crystal structure shows that the high affinity interaction between GP1 and hTfR1 forces the normally flexible loop in the apical domain of hTfR1 into a rigid $\beta$-pleated sheet domain. For GP1, several extended loops mediate binding to hTfR1 \citep{Abraham2010,Rad20112}, and many of the interface interactions are mediated by extensive hydrogen-bonding networks \citep{Abraham2010}. Experimental alanine-scanning and whole-cell infectivity assays have identified several sites in both GP1 and hTfR1 that are probably critical for establishing infection \citep{Rad20111,Rad20112}.

@@

\subsection{The GP1/hTfR1 system}
The GP1/hTfR1 interface (Figure~\ref{fig:networks}) marks a particularly important and useful test system. There are several sites on both the human and viral protein known to affect the infectivity phenotype of MACV. Many of the important sites have been mapped by \textit{in vitro} flow-cytometry based entry assays. The GP1/hTfR1 interface appears not to be dominated by one particular type of interaction (electrostatics, hydrogen-bonding, or van der Waals). In addition, much of the binding domain on hTfR1 is on a loop that is flexible prior to viral binding, but organizes to become a strand of a $\beta$-sheet on binding. As a result,  many other computational techniques \citep{Grant2011,Kortemme2004} are only marginally useful. The complex nature of this interface represents a particularly difficult challenge for traditional computational analysis.

In total, we tested 7 point mutants and 3 double mutants in addition to the WT complex (Table~\ref{tab:summary_mutants}). All of the mutations are within 5~\AA\ of the protein--protein interface. Mutations in hTfR1 at site 211 have proven capable of causing loss-of-entry according to \textit{in vitro} flow-cytometry infection assays or known host-range limitations \citep{Rad2008,Rad20111,Rad20112}. Most likely, this effect is caused by the destruction of a critical hydrogen bond to Ser113 or Ser111 in GP1. The lost hydrogen bond would lead to the subsequent loss of a large hydrogen-bonding network seen in the crystal structure (Table~\ref{tab:summary_mutants}) \citep{Abraham2010}. In a manner similar to site 211, N348 appears to be important for binding by participating in a critical hydrogen bonding network \citep{Rad2008,Abraham2010} to GP1. In particular, N348Lys is reported in the literature to cause significantly reduced viral entry \textit{in vivo} (Table~\ref{tab:summary_mutants}) \citep{Rad2008,Abraham2010}. Finally, an alanine mutation at site 111 in GP1 (mutation vR111A) has also been shown to cause decreased entry (Table~\ref{tab:summary_mutants}) \citep{Rad20112}. For notation purposes, the viral site is always referred to with a preceding `v'.

@@

\subsection{Comparative analysis of the GP1/hTfR1 interface}
Considering the involvement of extended hydrogen-bonding networks in the GP1/hTfR1 interface (Figure~\ref{fig:networks}), it was not clear that individual alanine mutations, even those that should destroy such networks, would significantly change the strength of interaction. One major advantage of first principles simulations is the ability to test mutations other than alanine without additional underlying assumptions in the energy function. As shown in Table~\ref{tab:summary_mutants}, we made additional mutations based on biochemical intuition or available experimental data to chemically diverse amino acids including tryptophan, lysine, aspartate, and threonine. Several mutations caused significant relative affinity changes. In addition, to detect synergistic effects, we tested several double mutants where both mutations appeared to cause similar changes in binding. Then, we compared the size of those differences to single mutants (Figure~\ref{fig:sensitivity} and~\ref{fig:mutant_comparison}).
